---
title: "Lab_XX Assignment"
author: "Firstname Lastname, Section YY"
date: 'Due date'
output: 
  word_document:
    fig_caption: yes
    reference_docx: assignment_style_template.docx
---

```{r setup, include = FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(tigerstats)
lattice.options(default.theme = standard.theme(color = FALSE))
```


~ START 

Be sure to install and load the `tigerstats` library. Type the following code 
into the command console (if you haven't installed `tigerstats` already):

```
install.packages("tigerstats")
```  

Once `tigerstats` has been installed, delete this text block 
(from "~ START" to "~ END")

~ END


This template includes some settings:

* the header at the top includes a line that ensures figure captions are enabled
* the "reference_docx" line in the header at the top refers to the name of a Word document that provides the formatting style we wish to use; this file must be located in your working directory.  

You can replace all text below if you wish, but I would use the example R-chunks provided as a guide.


## Load required datasets

Use this R chunk to load any datasets that you will be analyzing

```{r}

```


Remember to use 2 number signs to precede question numbers, as follows:

## Question 1

You can also embed plots with figure captions, and with the output size specified, for example:

```{r fig.cap = "Figure 1: Scatterplot of sepal length vs sepal width for 150 iris plants from 3 species.", fig.width = 4, fig.height = 4}
plot(iris$Sepal.Length, iris$Sepal.Width)
```

You can play with the width and height settings to get a figure size that you like.

## Question 2

In Figure 1 we see a positive association between sepal length and width.
