---
title: "Lab_01 Assignment"
author: "Firstname Lastname, Section YY"
date: 'Due date'
output: 
  word_document:
    fig_caption: yes
    reference_docx: Lab_01_word_style_template.docx
---


```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(tigerstats)
lattice.options(default.theme = standard.theme(color = FALSE))
```

Be sure to install and load the `tigerstats` library before you start.

Type the following into the command console:
```
install.packages("tigerstats")  # if you haven't installed already
library(tigerstats)
```

You can replace all text below if you wish, but I would use the example R-chunks provided as a guide.

This template includes some new settings:

* the header at the top includes a line that ensures figure captions are enabled
* the "reference_docx" line in the header at the top refers to a Word document that provies the formatting style we wish to use

Remember to use 2 hashtags to precede question numbers, as follows:

## Question 1

You can also embed plots with figure captions, and with the output size specified, for example:

```{r fig.cap = "Figure 1: Scatterplot of sepal length vs sepal width for 150 iris plants from 3 species.", fig.width = 4, fig.height = 4}
plot(iris$Sepal.Length, iris$Sepal.Width)
```

You can play with the width and height settings to get a figure size that you like.

## Question 2

In Figure 1 we see a positive association between sepal length and width.
