---
title: "Practice Assignment"
author: "Firstname Lastname, Section YY"
date: "Due date"
output:
  pdf_document: default
  word_document:
    fig_caption: yes
    reference_docx: assignment_style_template.docx
---

```{r setup, include = FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(tigerstats)
lattice.options(default.theme = standard.theme(color = FALSE))
```
* * *

## Getting started

This is a practice assignment, provided to you in R Markdown format (a file with the "Rmd" extension in its name).  

Each assignment will be provided to you in this format, and will also be available in PDF format on Canvas, on the main Lab Materials module.  

You are to use the R Markdown file for completing your assignment, as instructed in the [Preparing and formatting assignments tutorial](https://people.ok.ubc.ca/jpither/modules/Preparing_formatting_assignments.html).

~ START 

Be sure to install and load the `tigerstats` library. Type the following code 
into the command console (if you haven't installed `tigerstats` already):

```
install.packages("tigerstats")
```  

Once `tigerstats` has been installed, delete this text block 
(the block from "~ START" to "~ END")

~ END


### This template includes some settings:

* the header at the top includes a line that ensures figure captions are enabled
* the "reference_docx" line in the header at the top refers to the name of a Word document that provides the formatting style we wish to use; this file must be located in your working directory. It is called "assignment_style_template.docx", and is available for download [here](https://people.ok.ubc.ca/jpither/more/assignment_style_template.docx).  

You can replace all text below if you wish, but I would use the example R-chunks provided as a guide.

* * *

## Load required datasets

Typically one of the first steps do doing an assignment will be to download a dataset.  

Here is the first R "chunk" that contains code to download a dataset and store it in an object (dataframe) called "students"

```{r}
students <- read.csv(url("https://people.ok.ubc.ca/jpither/datasets/students.csv"), header = TRUE)
```

Next you'll have some questions to answer for your assignment. **Don't** accidentally delete or modify the questions from this document!  

* * *

## Questions

Use three number signs (hashtags) to identify the question number:

### Question 1:

How many observations (rows) are in the `students` dataframe?

Type your answer in full sentences here, and describe what you did to get the answer.  

The R chunk below is where you'll put the commands required to find the answer. 

```{r}

```


### Question 2:

How many variables (columns) are in the `students` dataframe?  

Answer here.

```{r}

```

* * *

## Knit and submit your assignment

Now you can try "knitting" your assignment to a Word document (you may be lucky and be able to knit directly to PDF... give it a try).  

Follow the instructions in the tutorial!  